---
title: "CUNY 607"
subtitle: "Week 12 HW"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

<br>

__In this exercise we will migrate a database from MySQL to MongoDB.__

<br>

###Load libraries

```{r libraries, message=FALSE}

library(mongolite)
library(dplyr)
library(tidyr)
library(rvest)
library(ggplot2)
library(knitr)
```

<br>

###Preparation

Make sure you have MySQL installed. Then please run the following sql script in order to create the movie_review database:

[Github_Link](https://raw.githubusercontent.com/mehtablocker/cuny_607/master/cuny_607_week_2_hw.sql)

<br>

###Loading the SQL tables into R

Be sure you use the correct username and password for your particular MySQL setup! We then loop through all of the tables in the db and store them each as a data frame in one list called df_list.

```{r load_tables}
my_db <- src_mysql(dbname = 'movie_reviews', username = 'user', password = 'password')
my_tables <- db_list_tables(my_db$con)
print(my_tables)

df_list <- list()

for (i in my_tables){
  df_list[[i]] <- tbl(my_db, i) %>% collect(n=Inf)
}

df_list %>% head()
```

<br>

###Migrate to MongoDB

We then use the mongolite package in R to easily insert the data frames as collections into a Mongo database.  

Please be sure to set the variable `mongo_url` to your particular mongo database location. The default is "mongodb://localhost" but you may have a cloud version you prefer. For security purposes I saved my particular URL in a local file and access it with the `readLines()` function.  

You can also choose your own value for `db=` within the `mongo()` function. The default is "test".

```{r mongo_migrate}
mongo_url <- readLines("mongodb_server_url.txt")

for (i in my_tables){
  con <- mongo(collection = i, db = "test", url = mongo_url)
  con$insert(df_list[[i]])
  con$count()
  print(con$find() %>% head(10))
  con$drop()
}

```

We can see that the information in the collections is the same as the SQL tables.  

The advantages of using a NoSQL database are that it is more flexible in its structure and also takes up less disk space. This may be desirable if working with a very large dataset. However, the flexibility can be a disadvantage as well. Data relationships are not strictly enforced, which could lead to duplication and disorganization. Relational databases are often preferred if a dataset is not overly immense and / or cost is not an issue.
