---
title: "CUNY 607"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
subtitle: Final Project
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

<br>

###Options volatility

__We will now examine implied volatility in the market. We do this by acquiring data on ticker symbol SPY, which is an exchange-traded fund (ETF) that replicates the S&P 500 index. Using the underlying stock and options price data, we calculate and plot the volatility curve or "smile."__

<br>

###Load libraries

```{r libraries, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(rvest)
require(quantmod)
require(derivmkts)
library(knitr)
```

<br>

###Define variables

We set the ticker symbol to SPY and the interest rate to 2.25 percent.

```{r define_var}
interest_rate <- 0.0225
symbol <- "SPY"
```

<br>

###Get last closing price

We use the quantmod package to look up historical stock prices and save the most recent closing price.

```{r close_price, message=FALSE, warnings=FALSE}
price_df <- getSymbols(symbol, auto.assign=F, from=Sys.Date()-3) %>% as.data.frame()
names(price_df) <- names(price_df) %>% gsub(paste0(symbol, "\\."), "", .)
price_df <- price_df %>% 
  rownames_to_column(var="Date") %>% 
  mutate(Date=as.Date(Date)) %>% 
  arrange(desc(Date)) %>% slice(1)
stock_price <- price_df$Close
price_df %>% head() %>% kable()
```

<br>

###Get dividend info

We look up the empirical dividend stream and convert it into a yield percent (to plug into the options pricing model.)

```{r get_div, message=FALSE, warnings=FALSE}
div_df <- getDividends(symbol, from = Sys.Date()-365, to=Sys.Date()) %>% as.data.frame()
names(div_df) <- names(div_df) %>% gsub(paste0(symbol, "\\."), "", .)
div_df <- div_df %>% rownames_to_column(var="Date") %>% mutate(Date=as.Date(Date))
div_yield <- suppressWarnings( tryCatch( sum(div_df$div)/stock_price, error=function(e) 0))
div_yield
```

<br>

###Get options chain

We use the quantmod package to retrieve options prices from Yahoo Finance. The implied volatility values are often missing or incorrect, so we will calculate them ourselves using the Black-Scholes formula from the derivmkts package. We do that by looping through the list of options and using the necessary information as inputs - stock price, strike, option type and price, time to expiration, dividend yield, and interest rate.

```{r get_options, results="hide", message=FALSE, warnings=FALSE}
### Get option prices
option_chain_list <- getOptionChain(Symbols = symbol, Exp = NULL)
option_chain_df <- data.frame()

for (i in 1:length(option_chain_list)){
  if (!is.null(option_chain_list[[i]]$calls) & !is.null(option_chain_list[[i]]$puts)){

    exp <- names(option_chain_list)[[i]] %>% as.Date(., format="%b.%d.%Y")
    t <- (as.numeric(exp-Sys.Date()))/365
    call_df <- option_chain_list[[i]]$calls %>% mutate(callput="call", time=t, exp=exp)
    put_df <- option_chain_list[[i]]$puts %>% mutate(callput="put", time=t, exp=exp)

    ### Calculate Implied Vol
    bidIV <- vector()
    askIV <- vector()
    for (k in 1:nrow(call_df)){
      bidIV[k] <- tryCatch(bscallimpvol(s=stock_price, k=call_df$Strike[k], r=interest_rate, tt=call_df$time[k],
                                        d=div_yield, price=call_df$Bid[k]), error=function(e) NA)
      askIV[k] <- tryCatch(bscallimpvol(s=stock_price, k=call_df$Strike[k], r=interest_rate, tt=call_df$time[k],
                                        d=div_yield, price=call_df$Ask[k]), error=function(e) NA)
    }
    call_df <- call_df %>% mutate(bidIV=suppressWarnings(as.numeric(bidIV)),
                                  askIV=suppressWarnings(as.numeric(askIV)), midIV=(bidIV+askIV)/2)
    bidIV <- vector()
    askIV <- vector()
    for (k in 1:nrow(put_df)){
      bidIV[k] <- tryCatch(bsputimpvol(s=stock_price, k=put_df$Strike[k], r=interest_rate, tt=put_df$time[k],
                                       d=div_yield, price=put_df$Bid[k]), error=function(e) NA)
      askIV[k] <- tryCatch(bsputimpvol(s=stock_price, k=put_df$Strike[k], r=interest_rate, tt=put_df$time[k],
                                       d=div_yield, price=put_df$Ask[k]), error=function(e) NA)
    }
    put_df <- put_df %>% mutate(bidIV=suppressWarnings(as.numeric(bidIV)),
                                askIV=suppressWarnings(as.numeric(askIV)), midIV=(bidIV+askIV)/2)

    option_chain_df <- rbind.fill(option_chain_df, call_df, put_df)
  }
}
```

```{r display_option_df}
option_chain_df <- option_chain_df %>% 
  filter(!is.na(midIV)) %>% mutate(underlying=symbol,
                                   stock_price=stock_price,
                                   price_date = price_df$Date)
option_chain_df %>% head() %>% kable()
```

<br>

###Plot volatility curve

We use ggplot to show the volatility smile and term structure for all options expirations between three and nine months out.

```{r plot_vol}
option_chain_df %>% 
  filter(exp>Sys.Date()+90, 
         exp<Sys.Date()+270, 
         Strike>=stock_price & callput=="call" | Strike<stock_price & callput=="put", 
         OI>1000, 
         Bid>=0.25) %>% 
  mutate(Expiration=as.character(exp), midIV=midIV*100) %>% 
  ggplot(aes(x=Strike, y=midIV, colour=Expiration)) + geom_line() + labs(title=paste0(symbol, " Options Implied Volatility by Expiration and Strike"), x="Strike Price", y="Mid-Market Implied Volatility")
```
