---
title: "CUNY 607 HW Week 1"
subtitle: "Exploration of Mushroom Data Set"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Load the data

Use read.csv and readLines to get data directly from web

```{r load_data}
require(tidyr, quietly = T, warn.conflicts = F)
require(dplyr, quietly = T, warn.conflicts = F)

data_url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"
mushroom_df <- read.csv(data_url, header=F, stringsAsFactors=F)
attribute_url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names"
attribute_text <- readLines(attribute_url)
```


###Get column names

Use various regex and R commands to find, clean, and populate column names, thereby (mostly) avoiding manual labor.

```{r get_column_names}

headers_text <- attribute_text[106:140]
headers_list <- headers_text %>% strsplit(., "\\.")

column_names <- headers_list %>% unlist()
column_names <- column_names[column_names %>% grep(":", .)]
column_names <- column_names %>% strsplit(., ":") %>% lapply(., function(x) x[1]) %>% unlist() %>% trimws()
column_names[1] <- "edibility"
column_names <- column_names %>% gsub("-", "_", .) %>% gsub("\\?", "", .)
names(mushroom_df) <- column_names
```


###Replace abbreviations

Make the data frame more readable by replacing one-letter abbreviations with the full words. Do this by first wrangling the abbreviations into a list of data frames (where each list element corresponds to a column of mushroom_df), and then looping through the list to match the abbreviation with the full word to replace within mushroom_df.

```{r replace_abbreviations}

abbrev_dict <- paste(headers_text, sep="", collapse="") %>% strsplit(., ":") %>% unlist()
abbrev_dict <- abbrev_dict[-c(1,2)] %>% as.list()
abbrev_dict <- abbrev_dict %>% lapply(., function(x) strsplit(x, ",") %>% unlist() %>% trimws())
abbrev_dict <- abbrev_dict %>% lapply(., function(x) strsplit(x, " ") %>% unlist())
abbrev_dict <- abbrev_dict %>% lapply(., function(x) x[grep("=", x)])
abbrev_dict[[1]] <- abbrev_dict[[1]] %>% gsub(")", "", .)
abbrev_dict <- abbrev_dict %>% 
  lapply(., function(x) x %>% as.data.frame() %>% separate(., `.`, into=c("full", "short"), sep="="))

for (i in 1:ncol(mushroom_df)){
  mushroom_df[ ,i] <- abbrev_dict[[i]]$full[sapply(mushroom_df[ ,i], 
                                                   function(x) which(abbrev_dict[[i]]$short==x)) %>% unname()]
}
```

###View the cleaned-up data frame

This looks much nicer!

```{r exercise_6, results='asis'}
knitr::kable(mushroom_df %>% head() %>% select(1:10))
```
