---
title: "CUNY 605"
subtitle: "Week 7 HW Assignment"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

###Problem 1

The CDF of y = the minimum of n Uniform(a,b) _continuous_ random variables is:

$$
\begin{equation}
  F(y) = 1 - \Big(\dfrac{b-y}{b-a}\Big)^{n}
\end{equation}
$$
We are interested in the _discrete_ case of y = the minimum of n Uniform(1, k) discrete random variables.
The CDF is:

$$
\begin{equation}
  F(y) = 1 - \Big(1 - \dfrac{y}{k}\Big)^{n}
\end{equation}
$$


By recognizing that the above formula is an accumulation function, we can use it to get the Probability Mass Function of the discrete case:

$$
\begin{equation}
  f(y) = \Big(1 - \dfrac{y-1}{k}\Big)^{n} - \Big(1 - \dfrac{y}{k}\Big)^{n}
\end{equation}
$$

###Problem 2

_"Your organization owns a machine that has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every 10 years. What is the probability that the machine will fail after 8 years?"_

<br>

(a) __Geometric__

$$
\begin{equation}
  P(x>8) = \Big(1 - \dfrac{1}{10}\Big)^{8}
\end{equation}
$$
```{r geometric}
p <- 1/10
1-sum(dgeom(0:7, p))   #the R function starts at 0, so must do 0:7 instead of 1:8
1/p   #mean
sqrt((1-p)/p^2)   #sd
```

<br>

(b) __Exponential__

$$
\begin{equation}
  P(x>=8) = e^{-8/10}
\end{equation}
$$
```{r exponential}
lambda <- 1/10
pexp(8, lambda, lower.tail = F)
1/lambda   #mean
1/lambda   #sd
```

<br>

(c) __Binomial__

$$
\begin{equation}
  P(x=0) = \Big(\dfrac{9}{10}\Big)^{8}
\end{equation}
$$
```{r binomial}
p <- 1/10
n <- 8
dbinom(0, n, p)
n*p   #mean
sqrt(p*(1-p)*n)   #sd
```

<br>

(d) __Poisson__

$$
\begin{equation}
  PMF = \dfrac{\lambda^{k}e^{-\lambda}}{k!}
\end{equation}
$$
```{r poisson}
p <- 1/10
n <- 8
lambda <- n*p
dpois(0, lambda)
lambda   #mean
sqrt(lambda)   #sd
```
