---
title: "CUNY 607 Data Presentation"
subtitle: "A Simple Baseball Linear Model"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

<style type="text/css">
h3 {
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

  
###Libraries

First we load the necessary libraries

```{r library_load}

library(dplyr)
library(rvest)
library(ggplot2)
library(stringr)
library(knitr)

```

  
###Get the data

Then use rvest to scrape the MLB standings from espn.com

```{r get_data}

season <- 2018
url <- paste0('http://www.espn.com/mlb/standings/_/sort/winpercent/dir/desc/season/', season, '/group/overall')
url_html <- url %>% read_html()
html_tables <- url_html %>% html_table(fill=T)
teams <- html_tables[[2]]
teams %>% head()

# use regex to extract team acronym - get uppercase letters followed by lowercase, then delete last two characters
teams <- teams %>% rename(team=1) %>% mutate(team=str_extract_all(team, "[A-Z]{3,4}[a-z]") %>% substr(1, nchar(.)-2))
teams %>% head()

standings <- html_tables[[4]]
standings %>% head()

team_standings <- bind_cols(teams, standings)
team_standings %>% head() %>% kable()
```


###Build a simple linear model

Scoring runs should lead to winning

```{r build_model}

team_standings %>% ggplot(aes(x=RS, y=W)) + geom_point() + geom_smooth(method="lm", se=T)
basic_lm_fit <- lm(W ~ RS, data=team_standings)
summary(basic_lm_fit)

```

The R^2^ is around **0.69**, which is a Pearson correlation of about r=0.83. Not bad.


###Expand the model

Winning is also influenced by how many runs a team **allows**

```{r expand_model}

team_standings %>% ggplot(aes(x=DIFF, y=W)) + geom_point() + geom_smooth(method="lm", se=T)
multi_lm_fit <- lm(W ~ RS + RA, data=team_standings)
summary(multi_lm_fit)

```

The adjusted R^2^ has gone up to **0.90**. Adding in Runs Allowed made a significant improvement to the model.

###Summary

We acquired our data via web scraping and cleaned it using regular expressions. We then built a basic linear model, which was a decent fit. But we made it even better by expanding it slightly to a multivariate regression.  

We now have a simple but effective (and interpretable) model for handicapping how many Wins a baseball team can expect based on how many runs they score and how many runs they allow.
