---
title: "CUNY 607"
subtitle: "Week 2 HW Assignment"
author: "mehtablocker"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RMySQL)
library(dplyr)
library(dbplyr)

```

###Preparation

1. Make sure you have MySQL installed. Then please run the following sql script in order to create the movie_review database:

[Github_Link](https://raw.githubusercontent.com/mehtablocker/cuny_607/master/cuny_607_week_2_hw.sql)

2. You will need to have the following R packages installed and loaded:  RMySQL, dplyr, dbplyr


###Loading the SQL tables into R

Be sure you use the correct username and password for your particular MySQL setup!

```{r load_tables}
mr_db <- src_mysql(dbname = 'movie_reviews', username = 'user', password = 'password')

movie_df <- tbl(mr_db, "movie") %>% collect(n=Inf)
critic_df <- tbl(mr_db, "critic") %>% collect(n=Inf)
review_df <- tbl(mr_db, "review") %>% collect(n=Inf)

```


###Analyzing the data

1. First let's see all of the reviews, sorted by movie title and rating, along with the critic's name:

```{r join_tables, results="asis"}
review_enhanced_df <- review_df %>% 
  left_join(movie_df, by="movie_id") %>% 
  left_join(critic_df, by="critic_id") %>% 
  arrange(title, desc(movie_rating))

knitr::kable(review_enhanced_df)
```

2. Which movie had the highest average rating?

```{r best_movie, results="asis"}
average_ratings_df <- review_enhanced_df %>% 
  group_by(movie_id) %>% 
  summarise(movie_title=tail(title,1),
            avg_rating=mean(movie_rating, na.rm=T)) %>% 
  ungroup() %>% arrange(desc(avg_rating))
knitr::kable(average_ratings_df)
```

3. On average, what rating did each critic give across all movies?

```{r critic_rating, results="asis"}
critic_ratings_df <- review_enhanced_df %>% 
  group_by(critic_id) %>% 
  summarise(name=tail(critic_name,1),
            avg_rating=mean(movie_rating, na.rm=T)) %>% 
  ungroup() %>% arrange(desc(avg_rating))
knitr::kable(critic_ratings_df)
```

As expected, the spread of ratings by movie is wider than the spread by critic.

```{r boxplot}
boxplot(critic_ratings_df$avg_rating, average_ratings_df$avg_rating, names=c("Ratings, by Critic", "Ratings, by Movie"))
```
